---
title: "Project World Happiness Report"
author: "Alexandre Chouraki"
date: "17/10/2017"
output: 
      rmarkdown::html_vignette:
        css: css/CSS_Splendor/css/splendor.css
vignette: >
  %\VignetteIndexEntry{Project World Happiness Report}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

### What is the World Happiness Report Database ?

#### Context

The World Happiness Report is a landmark survey of the state of global happiness. The first report was published in 2012, the second in 2013, the third in 2015, and the fourth in the 2016 Update. The World Happiness 2017, which ranks 155 countries by their happiness levels, was released at the United Nations at an event celebrating International Day of Happiness on March 20th. The report continues to gain global recognition as governments, organizations and civil society increasingly use happiness indicators to inform their policy-making decisions. Leading experts across fields – economics, psychology, survey analysis, national statistics, health, public policy and more – describe how measurements of well-being can be used effectively to assess the progress of nations. The reports review the state of happiness in the world today and show how the new science of happiness explains personal and national variations in happiness.

#### Content

The happiness scores and rankings use data from the Gallup World Poll. The scores are based on answers to the main life evaluation question asked in the poll. This question, known as the Cantril ladder, asks respondents to think of a ladder with the best possible life for them being a 10 and the worst possible life being a 0 and to rate their own current lives on that scale. The scores are from nationally representative samples for the years 2013-2016 and use the Gallup weights to make the estimates representative. The columns following the happiness score estimate the extent to which each of six factors – economic production, social support, life expectancy, freedom, absence of corruption, and generosity – contribute to making life evaluations higher in each country than they are in Dystopia, a hypothetical country that has values equal to the world’s lowest national averages for each of the six factors. They have no impact on the total score reported for each country, but they do explain why some countries rank higher than others.

#### What do the columns succeeding the Happiness Score(like Family, Generosity, etc.) describe?

The following columns: GDP per Capita, Family, Life Expectancy, Freedom, Generosity, Trust Government Corruption describe the extent to which these factors contribute in evaluating the happiness in each country. The Dystopia Residual metric actually is the Dystopia Happiness Score(1.85) + the Residual value or the unexplained value for each country as stated in the previous answer.

**Before doing anything, I have to set up R, and load the data, which is in the data_raw file of the package**

```{r setup, include = TRUE}
knitr::opts_chunk$set(echo = TRUE, message = TRUE, eval = TRUE)
```

```{r libraries}
library(readxl)
library(utils)
library(assertthat)
library(dplyr)
library(tidyverse)
library(ProjetAlex)
library(dygraphs)
library(ggplot2)
```

*When building my package, I used the console to read the .csv files I downloaded on Kaggle, and converted it to .rda using devtools::use_data function. Now I am loading them.*

```{r loading the data}
load("../data/data2015.rda")
load("../data/data2016.rda")
load("../data/data2017.rda")

```


**Now that I have the data for each year, my first goal is to aggregate them in order to be able to observe the evolution of all countries over time**

```{r adding years to the data}
# first, I have to mutate the data in order to add a column with the year in each database
# I'm going to use a function I coded, in the R file, called add_year

data2015 <- add_year(data2015, 2015)
data2016 <- add_year(data2016, 2016)
data2017 <- add_year(data2017, 2017)

```

```{r uniformizing data}
# Now that I have the year of each observation, in order to bind all my data together, I have to tidy it so I have the same columns in each of my dataframes, and only vectors

data2015 <- data2015 %>% 
  select(-Region, -Standard.Error)
indx <- sapply(data2015 %>% 
                 select(-Country)
               , is.factor)
data2015[indx] <- lapply(data2015[indx], function(x) as.numeric(as.character(x)))

data2016 <- data2016 %>% 
  select(-Region, -Lower.Confidence.Interval, -Upper.Confidence.Interval)

indx <- sapply(data2016 %>% 
                 select(-Country)
               , is.factor)
data2016[indx] <- lapply(data2016[indx], function(x) as.numeric(as.character(x)))

data2017 <- data2017 %>% 
  select(-Whisker.high, -Whisker.low)
indx <- sapply(data2017 %>% 
                 select(-Country)
               , is.factor)
data2017[indx] <- lapply(data2017[indx], function(x) as.numeric(as.character(x)))

rm(indx)
```

```{r binding data}
#at long last, I can finally bind the data together
data_Happiness <- bind_rows(data2015, data2016, data2017)
```
** Let's start the analysis! **

Now, I can finally see how happiness has evolved in the last three years, for, say, top 5 countries in average happiness over the last 3 years.

```{r top happiness}
top_5 <- data_Happiness %>%
  group_by(Country) %>% 
  mutate(average_happiness = sum(Happiness.Score)/3) %>% 
  select(Country, average_happiness) %>% 
  distinct() %>% 
  head(5) %>% 
  pull(Country)

top_5

evolution_top_5 <- ggplot(data_Happiness %>% 
                             filter(Country %in% top_5),
                           aes(x = year, y = Happiness.Score, color = Country)) + geom_line() + theme_classic()   

evolution_top_5
```

The most surprising thing in this graph is the apparent drop of happiness in Canada.
How can we explain it? 

```{r Canada analysis over time}
#Here, we are going to use a separate function I coded which gives us the evolution of the main indicators for any given country

country_analysis("Canada")
```

This doesn't make much sense... The happiness in Canada dropped in a significant way from 2016 to 2017, but most indicators are either stable or increasing (especially Family, which should supposedly have the most impact on Happiness, along with GDP per Capita, which increases too). 

In hindsight, this is not such a surprise : all the indicators aren't actually scores of the countries, simply the extent to which this factors make Canada happier than the imaginary country of Dystopia. So this actually means that Family and Social Support accounts for Happiness in Canada more than in 2016. 

This is interesting, but not helpful at all for an analysis...
* I still don't know why Canada is less happy than it was.
* And apparently I have no way to figure it out, since looking for the evolution of the indicators is meaningless.

I guess the best way to analyze this data is actually to compare the extent to which one given factor contributes to Happiness in all countries. 

For example, let's see which country benefits the most from its GDP, from an happiness perspective. 

```{r}
    data_Happiness %>%
    filter(year == 2017) %>% 
    select(Country, Economy..GDP.per.Capita.) %>%
    distinct() %>%
    head(5)
    
```


