---
title: "Project World Happiness Report"
author: "Alexandre Chouraki"
date: "17/10/2017"
output: 
      rmarkdown::html_vignette:
        css: css/CSS_Splendor/css/splendor.css
params:
  country:
    label: "Name of the specific country you want to analyze"
    value: "France"
    input: text
vignette: >
  %\VignetteIndexEntry{Project World Happiness Report}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

### What is the World Happiness Report Database ?

#### Context

The World Happiness Report is a landmark survey of the state of global happiness. The first report was published in 2012, the second in 2013, the third in 2015, and the fourth in the 2016 Update. The World Happiness 2017, which ranks 155 countries by their happiness levels, was released at the United Nations at an event celebrating International Day of Happiness on March 20th. The report continues to gain global recognition as governments, organizations and civil society increasingly use happiness indicators to inform their policy-making decisions. Leading experts across fields – economics, psychology, survey analysis, national statistics, health, public policy and more – describe how measurements of well-being can be used effectively to assess the progress of nations. The reports review the state of happiness in the world today and show how the new science of happiness explains personal and national variations in happiness.

#### Content

The happiness scores and rankings use data from the Gallup World Poll. The scores are based on answers to the main life evaluation question asked in the poll. This question, known as the Cantril ladder, asks respondents to think of a ladder with the best possible life for them being a 10 and the worst possible life being a 0 and to rate their own current lives on that scale. The scores are from nationally representative samples for the years 2013-2016 and use the Gallup weights to make the estimates representative. The columns following the happiness score estimate the extent to which each of six factors – economic production, social support, life expectancy, freedom, absence of corruption, and generosity – contribute to making life evaluations higher in each country than they are in Dystopia, a hypothetical country that has values equal to the world’s lowest national averages for each of the six factors. They have no impact on the total score reported for each country, but they do explain why some countries rank higher than others.

#### What do the columns succeeding the Happiness Score(like Family, Generosity, etc.) describe?

The following columns: GDP per Capita, Family, Life Expectancy, Freedom, Generosity, Trust Government Corruption describe the extent to which these factors contribute in evaluating the happiness in each country. The Dystopia Residual metric actually is the Dystopia Happiness Score(1.85) + the Residual value or the unexplained value for each country as stated in the previous answer.

**Before doing anything, I have to set up R, and load the data, which is in the data_raw file of the package**

```{r setup, include = TRUE}
knitr::opts_chunk$set(echo = TRUE, message = TRUE, eval = TRUE)
```

```{r libraries}
library(readxl)
library(utils)
library(assertthat)
library(dplyr)
library(tidyverse)
library(ProjetAlex)
```

*When building my package, I used the console to read the .csv files I downloaded on Kaggle, and converted it to .rda using devtools::use_data function. Now I am loading them.*

```{r loading the data}
load("../data/data2015.rda")
load("../data/data2016.rda")
load("../data/data2017.rda")

```


**Now that I have the data for each year, my first goal is to aggregate them in order to be able to observe the evolution of all countries over time**

```{r concatenating data}
# first, I have to mutate the data in order to add a column with the year in each database
# I'm going to use a function I coded, in the R file, called add_year

data2015 <- add_year(data2015, 2015)
data2016 <- add_year(data2016, 2016)
data2017 <- add_year(data2017, 2017)

```

```{r}
# Now that I have the year of each observation, in order to bind all my data together, I have to tidy it so I have the same columns in each of my dataframes

data2015 <- data2015 %>% 
  select(-Region, -Standard.Error)
data2016 <- data2016 %>% 
  select(-Region, -Lower.Confidence.Interval, -Upper.Confidence.Interval)
data2017 <- data2017 %>% 
  select(-Whisker.high, -Whisker.low)

```
```{r}
#at long last, I can finally bind the data together
data_Happiness <- bind_rows(data2015, data2016, data2017)
```

** Now, I can finally plot the evolution of Happiness, and all other indicators, for any country I want **

```{r}

```





